### 2022-06-06 chapters/fairness/justicia/introduction.tex
upto theoretical analysis of justicia

\subsubsection{Fair Machine Learning} Statistical discrimination caused by classifiers has motivated researchers to formulate several definitions of fairness and develop associated algorithms to mitigate bias. In this chapter, we focus on a popular concept of fairness, known as group fairness.\change{Add causal fairness here}  Existing group fairness metrics mostly belong to three categories: \textit{independence}, \textit{separation}, and \textit{sufficiency}~\cite{mehrabi2019survey}. Independence metrics, such as demographic parity, statistical parity, and group parity, try and ensure the outcomes of a classifier to be independent of the groups that the individuals belong to~\cite{feldman2015certifying,dwork2012fairness}. Separation metrics\unsure{Incorrect definition}, such as equalized odds, define a classifer to be fair if the probability of getting the same outcomes for different groups are same, separately for data from each possible ground-truth label~\cite{hardt2016equality}. Sufficiency metrics\unsure{Check definition}, such as counterfactual fairness, constrain the probability of outcomes to be independent of individual's sensitive data given their identical non-sensitive data~\cite{kusner2017counterfactual}.

@@

Our contributions are summarised below:
\begin{itemize}
	\item We propose a unified SSAT-based approach, {\justicia}, to verify independence and separation metrics of fairness for different datasets and classifiers.
	%\item \blue{{\justicia} measures fairness metrics for pre-, in-, and post-processing \red{algorithm}s with respect to the data  generating distribution}.
	\item Unlike previously proposed formal probabilistic verifiers, namely FairSquare and VeriFair, {\justicia} verifies fairness for compound and non-Boolean sensitive features.%, and also \red{separation metrics}.
	\item Our experiments validate that our method is more accurate and scalable than the probabilistic verifiers, such as FairSquare and VeriFair, and more robust than the sample-based empirical verifiers, such as AIF360.
	\item We prove a finite-sample error bound on our estimated fairness metrics which is stronger than the existing asymptotic guarantees.
\end{itemize}

>>>>

\subsubsection{Fair Machine Learning} Statistical discrimination caused by classifiers has motivated researchers to formulate several definitions of fairness and develop associated algorithms to mitigate bias. In this chapter, we focus on a popular concept of fairness, known as group fairness. Existing group fairness metrics mostly belong to three categories: \textit{independence}, \textit{separation}, and \textit{sufficiency}~\cite{mehrabi2019survey}. Independence metrics, such as demographic parity, statistical parity, and group parity, try and ensure the outcomes of a classifier to be independent of the groups that the individuals belong to~\cite{feldman2015certifying,dwork2012fairness}. Separation metrics, such as equalized odds, define a classifer to be fair if the probability of getting the same outcomes for different groups are same, separately for data from each possible ground-truth label~\cite{hardt2016equality}. Sufficiency metrics\unsure{Check definition}, such as counterfactual fairness, constrain the probability of outcomes to be independent of individual's sensitive data given their identical non-sensitive data~\cite{kusner2017counterfactual}.

@@

Our contributions are summarised below:
\begin{itemize}
	\item We propose a unified SSAT-based approach, {\justicia}, to verify independence and separation metrics of group fairness and causal fairness metrics for different datasets and classifiers.
	%\item \blue{{\justicia} measures fairness metrics for pre-, in-, and post-processing \red{algorithm}s with respect to the data  generating distribution}.
	\item Unlike previously proposed formal probabilistic verifiers, namely FairSquare and VeriFair, {\justicia} verifies fairness for compound and non-Boolean sensitive features.%, and also \red{separation metrics}.
	\item Our experiments validate that our method is more accurate and scalable than the probabilistic verifiers, such as FairSquare and VeriFair, and more robust than the sample-based empirical verifiers, such as AIF360.
	\item We prove a finite-sample error bound on our estimated fairness metrics which is stronger than the existing asymptotic guarantees.
\end{itemize}
